Automated Formal Synthesis\\ of Safe Digital Controllers for Continuous Plants

### 2018-12-20 automatica/paper.tex
start of ifdef's for axelerator

\begin{abstract}                          % Abstract of not more than 200 words.
%\textcolor{red}{Cristina: Can you please write an initial version of the abstract.}
We present a sound and automated approach to synthesizing safe,
digital controllers for physical plants represented as linear,
time-invariant models. Models are defined as differential equations
with inputs, evolving over a continuous state space. The synthesis
accounts for errors caused by the digitization effects introduced by
the controller. Our approach uses counterexample-guided inductive
synthesis (CEGIS): an inductive generalisation phase produces a
possible solution that is known to stabilize the system but that may
not be safe for all initial conditions. Safety is then verified either
via BMC or abstract acceleration; if the verification step fails, a
counterexample is provided to the inductive generalisation and the
process iterates until a safe controller is obtained.  We demonstrate
the practical value of this approach by automatically synthesizing
safe controllers for physical plant models from the digital control
literature.
\end{abstract}

@@

In this paper, we examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.
However, the use of digital control architectures adds new challenges
caused by artefacts specific to digital control, such as the effects
of finite-precision arithmetic, time discretization and quantization
noise introduced by A/D and D/A conversion.
%
Given an LTI model, we develop an automatic technique for generating
correct-by-construction digital controllers that addresses all these
challenges. Specifically, we automatically synthesize stable and safe,
software-implemented embedded controllers.  
%Due to the complexity of
%such closed-loop systems, we focus on linear models with known
%configurations, and perform parametric synthesis of controllers.
%meeting a number of properties that define their desired behavior.

@@

Our work addresses challenging aspects of the control synthesis problem:  
we perform digital control synthesis over a ``hybrid model'',
namely a model encompassing a plant exhibiting continuous behavior, 
whereas the controller operates in discrete time and over a quantized domain.  
In particular, our model evaluates the effects of the quantizers (A/D and D/A
converters) and of time discretization, as well as representation
errors introduced by working  %% artifacts such as an
%% observer~\cite{astrom1997computer}
in a finite-precision domain. 
We reason about high level properties such as stability, robustness, and safety, which define the
behaviour of the whole model.  
Notably, safety requirements are frequently overlooked in conventional feedback control synthesis, but
play a key role in systems engineering, particularly for safety-critical applications of clear concern in modern autonomy contexts.

%% Using state of the art verifiers, we evaluate these models against
%% automatically synthesised controllers using CEGIS-based synsthesis that
%% incorporate the above mentioned artifacts in different models of
%% Single-Input/Single-Output (SISO) systems.
In this paper we present two novel approaches for the synthesis of
digital controllers, both of which make use of the framework known as counterexample guided
inductive synthesis (CEGIS)~\cite{jha-icse10,
  DBLP:conf/asplos/Solar-LezamaTBSS06}, 
  a technique from formal methods that has recently shown much promise and which we 
  export in this work to a control-theoretical setup.  
CEGIS is an iterative process, where each iteration performs inductive
generalisation based on counterexamples provided by a verification
oracle (see Section~\ref{ssec:cegis}). The inductive generalisation uses information 
about a limited number of inputs to compute a candidate solution
for the entire range of possible inputs. Our two instantiations of CEGIS are described next.

%% offer three modes of
%% operation that address different aspects of the problem.
%% {\em The first  approach} uses a transfer-function based analysis that
%% solves the problem of robust stability for closed loop systems for a
%% collection of unstable plants.  Inthis case, there is no safety
%% specification since we are working in the frequency domain.

\emph{The first approach} uses a multi-staged technique that starts by devising a
digital controller that stabilizes the plant model while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the model dynamics, considering the 
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the safety boundaries are not violated for any larger number of
iterations.  Requiring unfolding up to the completeness threshold, this
approach is computationally expensive.

Instead of unfolding the dynamics, \emph{the second approach}
employs \emph{abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the model simultaneously. 
Additionally, this approach uses the \emph{abstraction-refinement} template \cite{DBLP:conf/cav/ClarkeGJLV00},
enabling one to start with a very simple (abstract) description of the model 
regardless of the complexity of its dynamics, and to refine to more complex models only 
when a solution cannot be found.

@@

%% \addtodo{CD: the novelty compared to CAV'17 needs to be clarified}
This paper is an extension of~\cite{DBLP:conf/cav/AbateBCCDKKP17}
and makes the following contributions (the fourth and the fifth are
specific extensions over~\cite{DBLP:conf/cav/AbateBCCDKKP17}): 
%
\begin{enumerate}

@@

%%   \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
%%   function representations, which are inadequate to prove safety requirements. } 
%% %
\item We provide two novel algorithms: the first, a multi-staged approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We provide bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to models in discrete 
time, \note{in the current paper we present an approach for
  soundly evaluating safety in combined continuous-discrete time and space. [this is via the AA-based back-end. However this part is not in the experiments, afaik.]}

\item A further limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17} 
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers are employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic, and compare the two approaches.  \note{Additionally, we enhance the
  implementation with an SMT back-end, allowing us to also use reals in order
  to represent the states of the plant.}

\item \note{Shall we mention the broader set of benchmarks?}

@@

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants given as transfer functions by exploiting bit-accurate
verification of software-implemented digital controllers~\cite{Bessa16}. 
While this work also uses CEGIS, the approach is restricted to digital
controllers for stable closed-loop systems given as transfer function
models: this results in a static check on their coefficients.  By contrast,
in the current paper we consider a state-space representation of the
physical system, which requires ensuring the specification over actual
traces of the model, alongside the numerical soundness required by the
effects of discretisation and finite-precision errors.
\note{Next paragraph is naive for a control audience - amend} 
A~state-space model has known advantages over the transfer function
representation~\cite{Franklin15}, as 
% it naturally generalizes to multivariate systems (i.e., with multiple inputs and outputs); and 
it allows synthesis of controllers with guarantees on the internal dynamics, e.g., 
%to synthesize controllers that make the closed-loop model 
\emph{safety}.  Our work indeed focuses on the safety of internal states, 
which is by and large overlooked in the standard (digital) control literature, 
by default focussed on stability properties.  
Moreover, our work integrates an abstraction-refinement (CEGAR) step inside the main CEGIS loop.

@@

\textbf{Discretization Effects in Control Design --}
Classical work in control synthesis has often disregarded digitalization effects, 
\note{[this paragraph needs more work:]
whereas recent results in digital control have focused on separate aspects of discretization, 
including delayed response~\cite{Duggirala2015} and finite word length (FWL) semantics, 
with the goal either to verify~\cite{daes20161} or to optimize~\cite{oudjida2014design} given implementations. }

There are two different problems that arise from FWL semantics.  The first
is the error in the model dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding and
saturation errors during model manipulation.  
In~\cite{fialho1994stability}, a stability measure based on the error of the digital dynamics ensures that
the deviation introduced by FWL does not lead to instability.  
\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$-calculus to directly model the digital controller so that selected
parameters result by design in stable dynamics.  
The analyses in~\cite{DBLP:conf/hybrid/RouxJG15, DBLP:conf/hybrid/WangGRJF16} rely on an
invariant computation on the discrete dynamics using Semi-Definite
Programming (SDP):  while the former contribution uses bounded-input and bounded-output
(BIBO) properties to determine stability, the latter uses Lyapunov-based
quadratic invariants.  In both cases, the SDP solver uses floating-point
arithmetic and soundness is checked by bounding the error.  
An alternative is~\cite{park2016scalable}, where the verification of given control code is
performed against a known model by extracting an LTI model of the code by
symbolic execution: to account for rounding errors, an upper bound is
introduced in the verification phase.  
The work in~\cite{picasso2003stabilization} introduces invariant sets as a mechanism to
bound the quantization error effect on stabilization.  
% as an invariant set that always converges toward the controllable set.  
Similarly, 
\cite{liberzon2003hybrid} evaluates the quantization error dynamics and \note{[unclear] bounds its trajectory to a known region over a finite time period}.   
%This technique works for both linear and non-linear systems.

@@

Models \eqref{eq:ode1} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t)$ and  $u_k$, respectively. 
Feedback architectures can be employed to manipulate the properties and behaviors of the plant:    
we are interested in the synthesis of digital feedback controllers $u_k$,  
as practically implemented on Field-Programmable Gate Arrays or Digital Signal Processors. 
\note{next: why only this single issue, rather than all discretisation effects as presented earlier?} Later, we will address the issue of variable quantization, 
introduced by ADC/DAC conversion blocks of Figure \ref{fig:digitalsystem}, 
as classically studied in \cite{astrom1997computer}.

@@

\subsection{Symbolic execution}
We use the bounded model checker CBMC~\cite{ClarkeKL04} to model controller behaviour on digitial signals with finite
bit-precision.  We express these controller semantics as C programs, which is CBMC's preferred input language.  CBMC
symbolically executes every instruction in an input program by translating it to a boolean satisfiability problem.  At
the end of this process, the entire input program is translated to a boolean satisfiability formula which is satisfiable
iff a certain property about the program holds.  CBMC's API allows to specify these properties in the form of assertions
in the input program, and we use these properties to express that our generated controllers are modelled in a stable and
safe fashion. \note{Do we need more details here? Should we explain e.g. SSA or give examples of how assignments, loops,
etc. are turned into formulas? Part of Sec.~\ref{sec:errors} can probably be moved here.}

@@

We account for the presence of errors originating from a number of sources, shown in Figure~\ref{fig:digitalsystem}:
\begin{itemize}
\item We represent the plant variables $x(t)$ and $u(t)$, which are continuous quantities, with finite precision in our model. 
It is necessary to account for the errors introduced by this representation, 
which we do using standard interval arithmetic. 
\item The quantisation errors by the ADC, which converts the analogue signal $x(t)$ into the digital signal $x_k$, 
which is then fed into the controller cf. Figure~\ref{fig:digitalsystem}. 
We account for this error in the acceleration-based back-end by introducing a noise signal $\nu_1$ that over-approximates it. 
Further details are found in Section~\ref{sssec:abstraction}.  
In the multi-stage back-end we use bit-precise modelling in CBMC~\cite{ClarkeKL04}, a bounded model checker for C-programs, to accurately encompass this conversion. 
In other words, CBMC constructs a Boolean formula that explicitly unwinds the transition system for symbolically represented initial states. 
\note{[rephrase:] The possible bounds that this symbolically represented initial state may take are restricted by upper and lower bounds as described in section . } 
The Boolean formula constructed by CBMC models the ADC conversions precisely on the symbolically represented states at every time step. 
When we check for a safety property, the Boolean formula is satisfied only if there is an input state value for which the subsequent trace through the transition system eventually violates the safety property. 
\item The representation errors due to performing finite-precision arithmetic in the controller block. 
In the acceleration-based back-end, 
this is again treated by the introduction of a noise signal that bounds the error. 
The multi-stage back-end uses the bit-precise modelling of finite-precision arithmetic in CBMC, 
in the same way as the ADC conversion was handled. 
We assume that the ADC represents numbers with a higher precision than the controller \note{why is this reasonable?}, 
and thus the precision is limited by the controller. 
\item We assume that the DAC converts numbers to analogue signals with the same precision as the controller. 
As a result we can ignore any additional errors introduced by the DAC.  Both the assumptions about the ADC and the DAC are reasonable assumptions based on commonly available hardware. 
\end{itemize}

>>>>

\begin{abstract}                          % Abstract of not more than 200 words.
%\textcolor{red}{Cristina: Can you please write an initial version of the abstract.}
We present a sound and automated approach to synthesizing safe,
digital controllers for physical plants represented as linear,
time-invariant models. Models are defined as differential equations
with inputs, evolving over a continuous state space. The synthesis
accounts for errors caused by the digitization effects introduced by
the controller. Our approach uses counterexample-guided inductive
synthesis (CEGIS): an inductive generalisation phase produces a
possible solution that is known to stabilize the system but that may
not be safe for all initial conditions. 
\ifx\axelerator
Safety is then verified either
via Bounded Model Checking or abstract acceleration;
\else
Safety is then verified via Bounded Model Checking;
\fi 
if the verification step fails, a
counterexample is provided to the inductive generalisation and the
process iterates until a safe controller is obtained.  We demonstrate
the practical value of this approach by automatically synthesizing
safe controllers for physical plant models from the digital control
literature.
\end{abstract}

@@

In this paper, we examine the case of Linear Time Invariant (LTI)
models, for which the synthesis of controllers is well understood.
However, the use of digital control architectures adds new challenges
caused by artefacts specific to digital control, such as the effects
of finite-precision arithmetic, time discretization and quantization
errors introduced by A/D and D/A conversion.
%
Given an LTI model, we develop an automatic technique for generating
correct-by-construction digital controllers that addresses all these
challenges. Specifically, we automatically synthesize %stable and 
safe,
software-implemented embedded controllers.  
%Due to the complexity of
%such closed-loop systems, we focus on linear models with known
%configurations, and perform parametric synthesis of controllers.
%meeting a number of properties that define their desired behavior.

@@

Our work addresses challenging aspects of the control synthesis problem:  
we perform digital control synthesis over a ``hybrid model'',
namely a model encompassing a plant exhibiting continuous behavior, 
whereas the controller operates in discrete time and over a quantized domain.  
In particular, our model evaluates the effects of the quantizers (A/D and D/A
converters), as well as representation
errors introduced by working  %% artifacts such as an
%% observer~\cite{astrom1997computer}
in a finite-precision domain. 
We reason safety requirements, which are frequently overlooked in conventional feedback control synthesis, but
play a key role in systems engineering, particularly for safety-critical applications of clear concern in modern autonomy contexts.

%% Using state of the art verifiers, we evaluate these models against
%% automatically synthesised controllers using CEGIS-based synsthesis that
%% incorporate the above mentioned artifacts in different models of
%% Single-Input/Single-Output (SISO) systems.
\ifx\axelerator
In this paper we present two novel approaches for the synthesis of
digital controllers, both of which make 
\else
In this paper we present a novel approach for the synthesis of digital controllers,
making 
\fi
use of the framework known as counterexample guided
inductive synthesis (CEGIS)~\cite{jha-icse10,
  DBLP:conf/asplos/Solar-LezamaTBSS06}, 
  a technique from formal methods that has recently shown much promise and which we 
  export in this work to a control-theoretical setup.  
CEGIS is an iterative process, where each iteration performs inductive
generalisation based on counterexamples provided by a verification
oracle (see Section~\ref{ssec:cegis}). The inductive generalisation uses information 
about a limited number of inputs to compute a candidate solution
for the entire range of possible inputs. Our two instantiations of CEGIS are described next.

%% offer three modes of
%% operation that address different aspects of the problem.
%% {\em The first  approach} uses a transfer-function based analysis that
%% solves the problem of robust stability for closed loop systems for a
%% collection of unstable plants.  Inthis case, there is no safety
%% specification since we are working in the frequency domain.
\ifx\axelerator
\emph{The first approach} 
\else
Our approach
\fi
uses a multi-staged technique that starts by devising a
digital controller that stabilizes the plant model while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the model dynamics, considering the 
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the safety boundaries are not violated for any larger number of
iterations.

\ifx\axelerator
Instead of unfolding the dynamics, \emph{the second approach}
employs \emph{abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the model simultaneously. 
Additionally, this approach uses the \emph{abstraction-refinement} template \cite{DBLP:conf/cav/ClarkeGJLV00},
enabling one to start with a very simple (abstract) description of the model 
regardless of the complexity of its dynamics, and to refine to more complex models only 
when a solution cannot be found.
\fi

@@

%% \addtodo{CD: the novelty compared to CAV'17 needs to be clarified}
This paper is an extension of~\cite{DBLP:conf/cav/AbateBCCDKKP17}
and makes the following contributions (the fourth is a 
specific extension to~\cite{DBLP:conf/cav/AbateBCCDKKP17}): 
%
\begin{enumerate}

@@

%%   \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
%%   function representations, which are inadequate to prove safety requirements. } 
%% %
\ifx\axelerator
\item We provide two novel algorithms: the first, a multi-staged approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
\else  
\item We present a novel multi-staged approach to synthesising digital controllers, using on 
  an unfolding of the dynamics up to a completeness threshold, and bit precise Bounded Model
  Checking to account for sources of imprecision in the implementation of the digital control algorithm
  and the modelling of the plant.
  \fi

%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We provide bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

%\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to models in discrete 
%time, \note{in the current paper we present an approach for
%  soundly evaluating safety in combined continuous-discrete time and space. [this is via the AA-based back-end. However this part is not in the experiments, afaik.]}

\item A limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17} 
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers are employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic.  
  %\note{Additionally, we enhance the
  %implementation with an SMT back-end, allowing us to also use reals in order
  %to represent the states of the plant.}

@@

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants given as transfer functions by exploiting bit-accurate
verification of software-implemented digital controllers~\cite{Bessa16}. 
While this work also uses CEGIS, the approach is restricted to digital
controllers for stable closed-loop systems given as transfer function
models: this results in a static check on their coefficients.  By contrast,
in the current paper we consider a state-space representation of the
physical system, which requires ensuring the specification over actual
traces of the model, alongside the numerical soundness required by the
effects of discretisation and finite-precision errors.

A~state-space model has known advantages over the transfer function
representation~\cite{Franklin15}, as 
% it naturally generalizes to multivariate systems (i.e., with multiple inputs and outputs); and 
it allows synthesis of controllers with guarantees on the internal dynamics, e.g., 
%to synthesize controllers that make the closed-loop model 
\emph{safety}.  Our work indeed focuses on the safety of internal states, 
which is by and large overlooked in the standard (digital) control literature, 
by default focussed on stability properties.  
Moreover, our work integrates an abstraction-refinement (CEGAR) step inside the main CEGIS loop.

@@

\textbf{Discretization Effects in Control Design --}
Classical work in control synthesis has often disregarded digitalization effects.
Recent results in digital control, however, have focused on separate aspects of discretization, 
including delayed response~\cite{Duggirala2015} and finite word length (FWL) semantics, 
with the goal either to verify~\cite{daes20161} the correctness of the implementation or to optimize~\cite{oudjida2014design} the design.

There are two different problems that arise from FWL semantics.  The first
is the error caused by the inability to represent the exact
state of the physical system, while the second relates to rounding and
saturation errors during mathematical operations.  
In~\cite{fialho1994stability}, a stability measure based on the error of the digital dynamics ensures that
the deviation introduced by FWL does not lead to instability.  
\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$-calculus to directly model the digital controller so that selected
parameters result by design in stable dynamics.  
The analyses in~\cite{DBLP:conf/hybrid/RouxJG15, DBLP:conf/hybrid/WangGRJF16} rely on an
invariant computation on the discrete dynamics using Semi-Definite
Programming (SDP):  while the former contribution uses bounded-input and bounded-output
(BIBO) properties to determine stability, the latter uses Lyapunov-based
quadratic invariants.  In both cases, the SDP solver uses floating-point
arithmetic and soundness is checked by bounding the error.  
An alternative is~\cite{park2016scalable}, where the verification of given control code is
performed against a known model by extracting an LTI model of the code by
symbolic execution: to account for rounding errors, an upper bound is
introduced in the verification phase.  
The work in~\cite{picasso2003stabilization} introduces invariant sets as a mechanism to
bound the quantization error effect on stabilization.  
% as an invariant set that always converges toward the controllable set.  
Similarly, 
\cite{liberzon2003hybrid} evaluates the quantization error dynamics and \note{[unclear] bounds its trajectory to a known region over a finite time period}.   
%This technique works for both linear and non-linear systems.

@@

Models \eqref{eq:ode1} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t)$ and  $u_k$, respectively. 
Feedback architectures can be employed to manipulate the properties and behaviors of the plant:    
we are interested in the synthesis of digital feedback controllers $u_k$,  
as practically implemented on Field-Programmable Gate Arrays or Digital Signal Processors. 
%This requires us to consider the errors introduced by, we will address the issue of variable quantization, 
%introduced by ADC/DAC conversion blocks of Figure \ref{fig:digitalsystem}, 
%as classically studied in \cite{astrom1997computer}.

@@

\subsection{Symbolic execution}
We use the bounded model checker CBMC~\cite{ClarkeKL04} to model controller behaviour on digitial signals with finite
bit-precision.  We express these controller semantics as C programs, which is CBMC's preferred input language.  CBMC
symbolically executes every instruction in an input program by translating it to a boolean satisfiability problem.  At
the end of this process, the entire input program is translated to a boolean satisfiability formula which is satisfiable
iff a certain property about the program holds.  CBMC's API allows to specify these properties in the form of assertions
in the input program, and we use these properties to express that our generated controllers are modelled in a stable and
safe fashion. %\note{Do we need more details here? Should we explain e.g. SSA or give examples of how assignments, loops,
%etc. are turned into formulas? Part of Sec.~\ref{sec:errors} can probably be moved here.}

@@

We account for the presence of errors originating from a number of sources, shown in Figure~\ref{fig:digitalsystem}:
\begin{itemize}
\item We represent the plant variables $x(t)$ and $u(t)$, which are continuous quantities, with finite precision in our model. 
It is necessary to account for the errors introduced by this representation, 
which we do using standard interval arithmetic. 
\item The quantisation errors by the ADC, which converts the analogue signal $x(t)$ into the digital signal $x_k$, 
which is then fed into the controller cf. Figure~\ref{fig:digitalsystem}. 
\ifx\axelerator
We account for this error in the acceleration-based back-end by introducing a noise signal $\nu_1$ that over-approximates it. 
Further details are found in Section~\ref{sssec:abstraction}.  
In the multi-stage back-end we
\else
We
\fi
use bit-precise modelling in CBMC~\cite{ClarkeKL04}, a bounded model checker for C-programs, to accurately encompass this conversion. 
In other words, CBMC constructs a Boolean formula that explicitly unwinds the transition system for symbolically represented initial states. 
The values that the symbolically represented initial state may take are restricted by upper and lower bounds as described in section~\ref{}.
The Boolean formula constructed by CBMC models the ADC conversions precisely on the symbolically represented states at every time step. 
When we check for a safety property, the Boolean formula is satisfied only if there is an input state value for which the subsequent trace through the transition system eventually violates the safety property.

\item The representation errors due to performing finite-precision arithmetic in the controller block. 
\ifx\axelerator
In the acceleration-based back-end, 
this is again treated by the introduction of a noise signal that bounds the error. 
The multi-stage back-end uses 
\else
We use
\fi
the bit-precise modelling of finite-precision arithmetic in CBMC, 
in the same way as the ADC conversion was handled. 
We assume that the ADC represents numbers with a higher precision than the controller, 
and thus the precision is limited by the controller. This is a reasonable assumption based on commonly available hardware.
\item We assume that the DAC converts numbers to analogue signals with the same precision as the controller. 
As a result we can ignore any additional errors introduced by the DAC. Again, this assumption is reasonable based on commonly available hardware. 
\end{itemize}
